Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources

### 2019-01-17 paper.tex
Update with abstract and more indepth arguments.

\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{framed}
\usepackage{cleveref}

@@

\author{
\IEEEauthorblockN{Nicholas Hazekamp}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
nhazekam@nd.edu}
\and
\IEEEauthorblockN{Tim Shaffer}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
tshaffe1@nd.edu}
\and
\IEEEauthorblockN{Jakob Blomer}
\IEEEauthorblockN{\&}
\IEEEauthorblockN{P. Samuel M. Teuber}
\and
\IEEEauthorblockN{Douglas Thain}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
dthain@nd.edu}
}

@@

\begin{abstract}
High energy physics research at CERN has a large and rapidly growing need for computation.
To bridge the gap, may scientists are looking to HPC resources.
HPC environments provides a platform for large
batch computation, similar to the WLCG.
However, the restrictions on computation,
such as limited internet access and prohibiting FUSE,
clash with HEP's reliance on CVMFS as global filesystem
for software.
As most LHC jobs uses 
only a small portion of CVMFS,
a more fine-grained approach to preparing container images 
is needed to limit creation time and storage.
Shrinkwrap introduces a method for creating and
updating projections, but a framework for managing
theses is needed.
...
\end{abstract}

@@

High energy physics research at CERN has a large and rapidly growing need for computation.
With upgrades to the LHC scheduled in the coming years,
the amount of computation required will increase dramatically.
The simulation and analysis workloads at CERN are primarily performed using software distributed
via the CernVM File System (CVMFS).
CMVFS provides a read-only, distributed file system using HTTP for network transport and extensive local
caching to rapidly distribute updates in software across the WLCG.
The Worldwide LHC Computing Grid (WLCG) currently satisfies most of the computational needs of the LHC experiments.
With the upcoming high luminosity upgrades to the LHC,
the computational resources needed will nearly double (?? check this) by 2026,
% On CERN's page it estimate 50-100X need, argonne had a report that was 10-20X
prompting researchers to turn to available high performance computing (HPC) resources to meet demand.

HPC environments provides a platform for large
batch computation, similar to the WLCG.
In contrast with the WLCG's single slot submissions,
HPC applications typically consist of parallel jobs using several cores up to a full machine for each submission.
Most HPC systems have site-specific configuration and security requirements.
These can take the form of limited network availability at computation nodes,
limited container support,
and restricted access to privileged system operations (such as FUSE).
These restrictions present challenges when porting CERN's HEP applications,
such as difficulties submitting tasks,
inability to access required software,
and additional time spent converting between image and container formats.

Dependence on CVMFS to distribute experiment software currently makes it impossible to take advantage of many HPC sites.
CVMFS requires unrestricted network access and is implemented as a FUSE module,
which sites often disallow for security and management reasons.
Some sites, such as NERSC,
have attempted workarounds for providing CVMFS,
such as creating large images containing all of the software stored in CVMFS (500GB-1TB) or
maintaining separate specialize systems for distributing 
CVMFS, such as DVMFS.
However, both of these existing solutions require 
hours to days of attention from system administrators to deploy or update,
which leads to lag between the available software and the current version.
Additionally for systems that utilize large container images, large amounts of storage are needed to maintain different versions.

Since each LHC experiment job uses 
only a relatively small portion of CVMFS,
a more fine-grained approach to preparing container images would reduce the time spent creating images and allow for more precise control of the software versions in use.
This could also reduce the storage required at each worker node and the network bandwidth spent transferring images.
We modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of a live HEP applications.
Using this live tracing or static analysis,
we are able to determine exactly which parts of CVMFS must be included in a container image for a particular job.
We developed a tool, Shrinkwrap, that can \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
Such a \emph{projection} of CVMFS is suitable for inclusion in a static container image that does not depend on FUSE or network access.

When providing CMVFS at an HPC site,
the cost of creating, distributing, and storing customized images for a large number of jobs and users imposes additional challenges.
The na\"{i}ve approach of creating a projection of CVMFS for each job results in significant computation, storage, and bandwidth overhead.
Acknowledging that jobs frequently have significant overlap in dependencies,
we propose a metric for determining the similarity between projections.
This gives the possibility of identifying existing container images that could be re-used to fulfill a new submission.
In addition, we explore the trade-offs in automatically augmenting existing projections to allow container images to support multiple distinct job types.
We bring these tools together to propose a potential enhancement to a batch scheduler for making online decisions about whether to create, reuse, or augment projections of CVMFS to efficiently support high-throughput HEP jobs on HPC resources.

%TODO summarize results

@@

\section{File Specifications}

The first step in addressing the high footprint 
of CVMFS repositories is to provide accurate 
specifications of the files, libraries, and packages
needed for each experiment.
These file specifications can then be used when 
projecting CVMFS into realized filesystems.
In general, the ability to project global filesystems 
hinges on the clarity of the file specification.
File specifications can be derived from several
different sources, such as 
usage traces, static analysis, or package listings.
The end goal of any of these methods is to 
create a comprehensive list of needed files and structure.
The nature of global filesystems flexible behavior
further extends file specifications by requiring
not just the files and structure, but also what
version/revision was used to verify the contents as well.

@@

\subsection{Trace Sampling}
Tracing was selected as it provides a consistent 
behavior between experiments and use cases.
This allows the solution to port between applications
quickly without having to delve in to the applications
specific configuration.
As previously mentioned, there are several methods for
obtaining a trace from a file.

@@

\subsection{Defined Consistency}
In order for a file specification to accommodate 
the need for describing the structure and contents
the specification goes beyond a generic file listing.
A file specification needs to include enough information
to establish the directory structure and verify the contents.
This can be done in several ways, 
such as computing the checksum of contents,
using global revision numbers, 
or running verification software.
Computed checksums provide a system agnostic approach
to verification by allowing the checksum to be computed
at specification and verified before use.
Computed checksums provide a
consistency guarantee to the byte level,
but are often expensive at runtime for large
production systems.
This approach is alleviated in systems such as 
Git and CVMFS as the contents' checksums are 
computed at ingestion into the repository.
This method relies on the underlying system computing
and providing this information for verification.

@@

Aside from computing and checking the contents of the
repository or it's files, it is possible to verify the
software using experiment computation.
This is a common method used to testing software
and works equally when the target software provides
coverage of the features used by the application.
The difficulty here is in providing tests that
behave consistently, verify if computation is
incorrect, and cover the desired features.
Though it is possible for the submitting user
to provide test with coverage, application
developers have a better grasp of it usage
and how to provide coverage in testing.
An additional method is to run the application
on a determined site for initial results and
compare in the future for verification, but
this relies on deterministic code which is 
not always the assumption in HEP computing.

I AM NOT SURE WHICH WE WILL USE. WRAP THIS UP LATER.

@@

\subsection{Packing Projections}

All of the above methods describe utilize CVMFS
and rely on the internet connection to create
the projections.
A core assumption of this paper is that internet
access is limited at the compute nodes, so a method
is needed to transform and use these projections
for computation.
For this we looked at several methods for providing
these projections and compared using 
Container images (images specific to Docker, Singularity, ect.)
versus SquashFS images.
The goal is to find a flexible, compatible intermediate
format that allows projections to be moved between sites
and technologies.
For this purpose, DOcker images, Singularity images, and 
SqaushFS are compared on their time to create,
storage requirements, and most importantly compatibility
with several used technologies, namely 
Docker, Singularity, and Shifter.

@@

TABLE COMPARING COMPATIBILITY AND SUPPORT

%MERGING PROJECTIONS

%INCOMPATIBLE PROJECTIONS
%It is often the case when using CVMFS
%an analysis job requires pieces of several repositories.
%As a result a method for merging and verifying that these
%repositories are compatible is needed.

@@

\subsection{Use Cases}
Here we illustrate several real world use cases.
These use cases are being used as the existing solutions
were not feasible and a new approach was needed.
In analyzing these use cases we can see the beginnings of
the problem outlined in the previous subsection.

@@

A major queston not addressed by the Shrinkwrap tool is how to manage images after creation.
In the case of a small number of manually created container images,
this is not a serious problem.
Thus for integrating with the existing systems at NERSC,
Shrinkwrap alone might be sufficient.
We would also like to support CVMFS as a site-wide service,
e.g.\ with a batch system plugin that can automatically prepare container images on job submission.
This both minimizes direct administrator burden and more closely approximates user experience on other resources like the WLCG.
With static analysis and dynamic tracing of tasks,
we have ways to choose subsets of CVMFS to project to run specific tasks.
The user can collect this information themself,
or tracing, too, could be handled automatically.
We have seen two approaches to this in section ??.
If the same types of jobs are also run on other resources such as the WLCG,
it is possible to instrument a fraction of runs to build a profile of the type of task.
Alternatively, we could integrate tracing with image creation by automatically running exemplar tasks on the WLCG or on dedicated bridge nodes and collecting traces as necessary.
Assuming that accurate dependency information is available via some means,
we are interested in automatically providing suitable projections of CVMFS to users.

@@

The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
Unfortunately, this solution is very wasteful of computing time.
The average simulation or analysis job runs for XXX seconds,
while preparing an image for it using Shrinkwrap and mksquashfs takes on average YYY seconds.
This image creation is a serial step that must be completed before any actual processing occurs.
In addition, freshly created images must be transferred and stored on any nodes that will use them,
increasing load on the network and storage infrastructure.

The first step toward improving performance is to re-use previously created container images.
A single experiment submitting computational tasks is likely to request the same projection of CVMFS,
at least for jobs submitted around the same time.
It is straightforward to check for exact equality between projections as long as the system retains the set of files that went into their creation.
In particular, it is important to retain both the paths and the hash of the file contents or some other unique identifier.
This ensures that multiple versions of the same file are recognized as distinct entities.

Next, worker nodes need to be able to cache previously used container images.
Technologies like Docker currently operate this way (often to the chagrin of users and administrators doing scientific computing).
Since projections of CVMFS can be re-created on demand,
there is no need to carefully track which container images to keep.
For our evaluation,
we chose LRU as a simple caching policy.
Worker nodes are assumed to have some fixed amount of storage available for caching images.
Node-local scratch storage is sufficient for temporarily caching these images.
We also assumed that durable storage is available on the bridge machine for storing the data cache and projections associated with Shrinkwrap,
as well as a cache of container images derived from these projections.
Again, we chose LRU as a caching policy for our evaluation,
though administrators would be free to experiment with different policies.

@@

\subsection{Evaluation of the Initial Approach}

We wanted to evaluate the viability of the simple design discussed in the previous section,
so we analyzed of a stream of actual job submissions to the WLCG.
Figure ?? gives a summary of the results.
We observed that for a single experiment at a given point in time,
the requested projection is fairly stable.
However, updates to software sometimes overlapped with previous job submissions,
leading to multiple projected versions of repos at the same time.
In addition, submissions from different experiments generally request different versions of repos.
Thus for a site servicing only a single LHC experiment and with ample storage for caching images,
the simple approach outlined above is sufficient.
For a more diverse workload like that of the WLCG,
however, the amount of resources spent creating and storing projected images rapidly increases.

This presents a problem for sites attempting to significantly augment existing computational capacity,
as they must be able to efficiently support jobs that vary widely both across experiment and time.
Since NERSC was willing to devote QQ~TB of storage to complete copies of CVMFS for their current container images,
the storage cost may be bearable.
It is not safe to assume,
however, that this will generally be the case.
Thus we need to find a way to more efficiently handle and store images while still retaining the flexibility of fine-grained per-job customization.

It is important to note that while the cumulative size of the generated container images grows quickly,
the projections themselves take up much less space.
This is due to the deduplication feature of the Shrinkwrap tool.
When creating a projection,
Shrinkwrap fetches any missing file data and stores it in a common content-addressable cache directory.
Shrinkwrap then hard links into the data cache to create the projection.
This allows multiple projections with contents in common to share their underlying storage.
Thus even with a large number of projections,
as long as the differences between projections are small the storage space required is modest.
When creating container images,
however, this deduplication is broken as it is not possible to share components between disk images.
When creating each image all data in the projection must be copied in.
We observed that the overlap between images was in many cases quite large.
We therefore investigated strategies for deduplication between static disk images,
as this would greatly reduce the storage and transfer overhead observed here.

@@

To decide if two projections are ``close'',
we define the parameter $\alpha$ as the maximal Jaccard distance between closely related projections.
Since Jaccard distance is by definition between zero and one,
$\alpha$ must be in the same range.
Choice of $\alpha$ is left to the system administrators.
Very small choices of $\alpha$ require that projections be extremely close before considering them for merging.
In the extreme case with $\alpha = 0$,
only \emph{identical} images will be considered close so no images will be merged.
This results in a larger number of independent images.
Choosing $\alpha$ to be larger makes it more likely for images to be considered similar and merged.
This results in more augmented images that serve multiple tasks.
In the extreme case of $\alpha = 1$,
\emph{every} pair of images is considered close and merged if possible.
This results in large container images that have accumulated many projections.
We explore tradeoffs when choosing $\alpha$ and give recommendations based on observed patterns of activity in section ??.

>>>>

\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{framed}
\usepackage{cleveref}
\usepackage{hyperref}

@@

\author{
\IEEEauthorblockN{Tim Shaffer}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
tshaffe1@nd.edu}
\and
\IEEEauthorblockN{Nicholas Hazekamp}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
nhazekam@nd.edu}
\and
\IEEEauthorblockN{Jakob Blomer}
\and
\IEEEauthorblockN{Douglas Thain}
\IEEEauthorblockA{University of Notre Dame \\
Notre Dame, Indiana 46556 \\
dthain@nd.edu}
}

@@

\begin{abstract}
Scientific applications often depend on large, complex software repositories containing libraries, tools, and frameworks in multiple versions and platforms to support many researchers.
As a practical necessity, software repos are most commonly distributed via shared or global filesystems.
In moving existing research to HPC environments, however,
scientists must work around restrictions that can interfere with global software delivery methods.
To illustrate these challenges,
we examine CVMFS, the CernVM File System.
CVMFS is a global, distributed filesystem that provides read-only access to experiment software and configurations for high energy physics research at CERN.
CVMFS requires a number of network and OS configurations that are incompatible with a restrictive HPC environment.
In this work, we introduce a set of tools and procedures for \emph{projecting} the global filesystem (or parts thereof) into local container images suitable for HPC sites.
As part of our evaluation,
we found that using local approximations of a global filesystem tends to introduce problems of its own in terms of compute and storage overhead.
We thus examined several declarative and profiling-based approaches to creating more fine-grained projections to fit application requirements,
and propose a strategy for efficiently managing these projected container images at an HPC site.
Using our projection tools and image management strategy,
we implemented a proof-of-concept submission system for running existing HEP analysis tasks in an HPC environment.
\end{abstract}

@@

Scientific research has a large and rapidly growing need for computation.
Researchers have a number of options at their disposal to meet those needs,
including campus-local or organization-local resources,
cloud providers,
grid computing,
and large HPC sites.
Moving applications between these classes of resources can present problems,
as the assumptions made during development may no longer hold.
For example, organization-local resources and some grid sites commonly provide access to shared filesystems at each compute node.
In addition, there is some flexibility in system configuration to support an organization's primary research activity.
A new computing environment may be much less open to changes for a specific application,
especially large, multi-tenant resources such as HPC sites.
In these cases, it is left to scientists to adapt their application software to the available resources.

Many of our research collaborators have experienced significant frustration in preparing an existing scientific application to run on a new type of resource.
As a concrete example for this work,
we consider the challenges in running simulation and analysis workloads at CERN on HPC sites.
These applications assume several pieces of infrastructure to be in place and require some configuration and customization to the compute environment.
This is not an issue when running on Worldwide LHC Computing Grid~(WLCG),
which is a dedicated resource built to meet CERN's substantial computing needs.
The WLCG currently satisfies most of the computational needs of the LHC experiments at CERN.
With the upcoming high luminosity upgrades to the LHC,
the computational resources needed will will increase by at least an order of magnitude by 2026,
% On CERN's page it estimate 50-100X need, argonne had a report that was 10-20X
prompting researchers to turn to available high performance computing (HPC) resources to meet demand.
%In contrast with the WLCG's single slot submissions,
%HPC applications typically consist of parallel jobs using several cores up to a full machine for each submission.

Most HPC systems, however, have site-specific configuration and security requirements.
These can take the form of limited network availability at computation nodes,
limited container support,
and restricted access to privileged system features (such as FUSE support).
These new restrictions present challenges when porting existing scientific applications,
which expect required software to be readily available,
make a number of assumptions about how the system should be configured,
and may not be designed with tasks such as converting between image and container formats in mind.

In the case of the simulation and analysis workloads at CERN,
unrestricted network access and FUSE support are critical for providing global filesystem access.
Nearly all research software used at CERN assumes the availability of the CernVM File System (CVMFS),
a distributed global filesystem built for distributing software and configuration data.
CMVFS provides a read-only, distributed file system using HTTP for network transport and extensive local
caching to rapidly distribute updates in software across the WLCG.
Dependence on CVMFS to distribute experiment software currently makes it impossible to take advantage of many HPC sites.
CVMFS requires unrestricted network access and is implemented as a FUSE module,
both of which sites often disallow for security and management reasons.

Some sites, such as NERSC,
have attempted workarounds for providing CVMFS,
such as creating large images several terabytes in size that contain all of the software stored in CVMFS or
maintaining local mirrors of CVMFS available via site-local shared filesystems.
However, both of these existing solutions require 
hours to days of attention from system administrators to deploy or update,
which leads to lag between the available software and the current version.
Additionally for systems that utilize large container images, large amounts of storage are needed to maintain different versions.

Since each LHC experiment job uses 
only a relatively small portion of CVMFS,
a more fine-grained approach to preparing container images would reduce the time spent creating images and allow for more precise control of the software versions in use.
This could also reduce the storage required at each worker node and the network bandwidth spent transferring images.
It is often possible to determine dependencies for particular applications statically by examining submission commands, setup scripts, etc.
This approach, while reliable, tends to be labor-intensive and tied to specific experiments and software stacks.
We therefore modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of a live HEP applications.
Using live tracing and static analysis,
we are able to determine exactly which parts of CVMFS must be included in a container image for a particular job.
We developed a tool, Shrinkwrap, that can efficiently \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
Such a \emph{projection} of CVMFS is suitable for inclusion in a static container image that does not depend on FUSE or network access.

While working with projected container images and Shrinkwrap,
we found that the cost of creating, distributing, and storing customized images for a large number of jobs and users creates unexpected challenges as well.
Shrinkwrap is designed to efficiently manage projections by deduplicating data in the filesystem.
Disk images, however, must contain complete copies of all included data.
The na\"{i}ve approach of creating a projection of CVMFS for each application results in frequent duplication of data and significant overhead in terms of computation and bandwidth.
Depending on the granularity of the projection,
image creation time can easily exceed individual job's actual compute time.
The alternative extreme, i.e.\ building a single large image that satisfies the requirements of all applications,
forces all jobs to use the same repository revision,
which can result in incompatibilities across applications.
Updates to the repository or changes in applications and requirements can again result in large amounts of overhead due to frequently creating, transferring, and storing these very large images.

Observing that jobs frequently have significant overlap in dependencies,
we designed a system to more efficiently manage container images on job submission.
First, we defined a metric for determining the similarity between projections.
This gives the possibility of identifying existing container images that could be re-used to fulfill a new request.
We also determined that in some cases,
it is less expensive to augment an existing, similar projection to allow a single container image to support multiple distinct applications.
We bring these tools together as a proof-of-concept submission system for making online decisions about whether to create, reuse, or augment container images to efficiently support high-throughput HEP jobs on HPC resources.

Compared to existing solutions,
Shrinkwrap created more compact projections than traditional methods such as rsync copies,
and allowed for more efficient updates to existing projections.
Based on the existing projects to produce static CVMFS container images and traces of benchmark applications,
we examined the sizes and structures of several projected container images.
Based on these properties,
we simulated the behavior of our image management system over a large number of job submissions.
We found that our approach can effectively manage the container images for a continuous stream of submissions with varying software dependencies and versions with less storage overhead than creating a new image per application.
In addition, our results show that applications can still take advantage of precise dependency and version specifications to reduce image creation and transfer overhead compared to working with large/complete mirrored copies of CVMFS.

@@

\section{Defining Application Dependencies}

While the entirety of CVMFS is fairly large,
a single application would never use every piece of data.
Rather than attempting to provide a complete copy of CVMFS locally,
which entails significant storage costs,
we could instead provide a limited subset of the complete filesystem that is sufficient for a particular job.
This approach is more in line with traditional software packaging practices,
where software is provided as needed in the form of RPM files or another similar format.
Working with software packages is simpler in that dependencies are explicitly specified in the package file or database.
With CVMFS, no such definitive specification exists.
This issue is common to shared/global filesystems,
since users expect the complete contents to be available on demand and so devote little time to explicitly defining dependencies and other requirements.

In the case of CVMFS,
building an accurate 
specification of the files, libraries, and packages
needed for each experiment is the first step toward creating usable container images.
Such file specifications can be used when 
projecting CVMFS into realized filesystems.
The correctness and storage overhead of the projected filesystem hinges on the clarity and accuracy of the file specification.
File specifications can be derived from several
different sources, such as 
usage traces, static analysis, or package listings.
The end goal of any of these methods is to 
create a comprehensive list of needed files and structure.
The nature of global filesystems flexible behavior
further extends file specifications by requiring
not just the files and structure, but also what
version/revision was used to verify the contents as well.

@@

\subsection{Trace Sampling}
Tracing is attractive as it provides a consistent 
behavior between experiments and use cases.
In addition, it is easier to automate and sample arbitrary applications without assistance from each user.
This allows the solution to port between applications
quickly without having to delve in to the applications
specific configuration.
As previously mentioned, there are several methods for
obtaining a trace from a file.

@@

\subsection{Unique Consistency Requirements}
CVMFS differs from traditional POSIX filesystems in that previous versions of all content remain available.
A researcher may specify the precise revision under which an application is known to run.
Thus the same path may refer to two completely different files for a pair of applications.
This is beneficial from the standpoint of reliability and reproducibility,
but complicates the process of projection.
A file specification needs to include enough information
to establish the directory structure and verify the contents at a specific revision.
This can be done in several ways, 
such as computing the checksum of contents,
using global revision numbers, 
or running verification software.
Computed checksums provide a system agnostic approach
to verification by allowing the checksum to be computed
at specification and verified before use.
Computed checksums provide a
consistency guarantee to the byte level,
but are often expensive at runtime for large
production systems.
This approach is alleviated in systems such as 
Git and CVMFS as the contents' checksums are 
computed at ingestion into the repository.
This method relies on the underlying system computing
and providing this information for verification.

@@

Aside from computing and checking the contents of the
repository or its files, it is possible to verify the
software using experiment computation.
This is a common method used to testing software
and works equally when the target software provides
coverage of the features used by the application.
The difficulty here is in providing tests that
behave consistently, verify if computation is
incorrect, and cover the desired features.
Though it is possible for the submitting user
to provide test with coverage, application
developers have a better grasp of it usage
and how to provide coverage in testing.
An additional method is to run the application
on a determined site for initial results and
compare in the future for verification, but
this relies on deterministic code which is 
not always the assumption in HEP computing.

\begin{table*}[t]
\begin{center}
\begin{tabular}{|c|r r r|r r r|}
\hline
CVMFS & & ls -lr & & & Shrinkwrap & \\
Repository & Total Size (B) & inode Count & Time to walk & 
  Data Files & Deduped Data Files & Dedup Bytes \\ \hline
%alice-ocdb & 482,642,307,428 & 1,621,661 & 2:22 &
% 1,621,282 & 1,595 & 24,433,131,203\\
alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
  797,404 & 23,996,541 & 2,098,509,015,172\\
%atlas-condb & 719,548,941,240 & 11,781 & 0:04 &
%  10,940 & 229 & 471,492,086 \\
atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
boss & 120,048,527,315 & 2,418,025 & 4:18 &
 256,061 & 1,868,192 & 72,301,955,666\\
cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
 18,667,178 & 101,448,867 & *2,242,658,825,701\\
cvmfs-config & 16,546 & 9 & 0:00 &
 2 & 0 & 0 \\
geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
 323,280 & 6,219,891 & 68,381,969,660 \\
grid & 30,642,826,259 & 714,235 & 1:17 &
 141,622 & 583,923 & 18,804,003,040 \\
lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
 10,032,339 & 20,406,235 & 368,070,277,634 \\
na61 & 5,222,967,734 & 25,602 & 0:02 &
 8,609 & 15,001 & 2,425,507,431 \\
sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
 \hline
\end{tabular}
\caption{CVMFS repository sizes.}
\label{tab:repo-sizes}
This table shows the size, number of inodes, and the amount
of duplicate data that is eliminated using CVMFS.
\end{center}
\end{table*}

\if 0
I AM NOT SURE WHICH WE WILL USE. WRAP THIS UP LATER.
\fi

@@

\subsection{Packaging Projections}

All of the above methods describe utilize CVMFS
and rely on the internet connection to create
the projections.
A core assumption of this paper is that internet
access is limited at the compute nodes, so a method
is needed to transform and use these projections
for computation.
For this we looked at several methods for providing
these projections and compared using 
Container images (images specific to Docker, Singularity, ect.)
versus SquashFS images.
The goal is to find a flexible, compatible intermediate
format that allows projections to be moved between sites
and technologies.
For this purpose, Docker images, Singularity images, and 
SqaushFS are compared on their time to create,
storage requirements, and most importantly compatibility
with several used technologies, namely 
Docker, Singularity, and Shifter.

@@

@@

\subsection{Use Cases of Shrinkwrap}
%Here we illustrate several real world use cases.
%These use cases are being used as the existing solutions
%were not feasible and a new approach was needed.
%In analyzing these use cases we can see the beginnings of
%the problem outlined in the previous subsection.

@@

With static analysis and dynamic tracing of tasks,
we have methods to choose subsets of CVMFS to project to run specific tasks.
Users can collect this information themselves,
or tracing, too, could be handled automatically.
We have seen two approaches to this in section ??.
For applications already running on the WLCG,
it is also possible to instrument a fraction of runs to build a profile of the type of task.
Alternatively, we could integrate tracing with image creation by automatically running a small number of sample tasks on the WLCG or on dedicated bridge nodes and collecting traces as necessary.
These approaches give a fair amount of flexibility in determining accurate dependency information.
Based on this information,
we are interested in automatically providing suitable projections of CVMFS to users.

Shrinkwrap provides a mechinsm for projecting parts of CVMFS,
but is not by itself a complete solution as it does not integrate with job submission or manage images after creation.
In the case of a small number of manually created container images,
this is not a serious problem.
Administrators could manage a small collection of images in this case and adjust jobs to work with their setup.
Thus for integrating with the existing systems at NERSC,
Shrinkwrap alone might be sufficient.
To support CVMFS as a site-wide service with a potentially varied user base and workload,
a more automated solution
(e.g.\ a batch system plugin that can automatically prepare container images on job submission)
is required.
This both minimizes direct administrator burden and more closely approximates user experience on other resources like the WLCG.
Our first attempt at such a system,
described in the next sections,
took the simplest possible approach to image.
As we were developing this approach, however,
a number of issues in real-world usage led us to conculde that a somewhat more sophisticated solution was required.

@@

The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
Unfortunately, this solution is very wasteful of computing time and storage space.
Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens to hundreds of gigabytes,
depending on the applicaion.
The full-repo images being tested at NERSC are multiple terabytes in size and take the majority of a day to create.
In addition, image creation must be completed before any actual processing occurs,
i.e.\ there is no way to start with a partial image that supports the first steps of a pipeline and finish building the image as the pipeline runs.
There is also the cost of transferring and storing freshly created images on any nodes that will use them,
further delaying job dispatch and increasing load on the network and storage infrastructure.

The disk I/O alone required to create even a targeted image is likely to take more time than a single simulation or analysis job.
Thus a simple performance improvement is to re-use previously created container images for multiple jobs.
A single experiment submitting computational tasks is likely to request the same projection of CVMFS,
at least for jobs submitted around the same time.
It is straightforward to check for exact equality between projections as long as the system retains the set of files that went into their creation.
Note that it is important to retain both the paths and the hash of the file contents used in images.
This ensures that multiple versions of the same file over time are recognized as distinct entities.
If there are few CVMFS users at a site,
this method could be sufficient.

After creating a set of static images,
the next step is to transfer them to local caches on the workers.
Worker nodes are assumed to have some fixed amount of storage available for caching images.
If the size of the ``working set'' of images is smaller than the node-local storage,
then each worker can keep a complete set of images.
While ideal, this case is not necessarily realistic beyond a few images.
With small/medium images in the tens-to-hundreds of gigabytes range,
maintining a complete set of projected CVMFS images at each worker requires substantial resource provisioning.
Thus it is important for workers be able to to cache only a subset of the available projected images.
There are numerous caching policies and options available,
but choice of worker cache strategy is not critically important for this application.
Since projections and images of CVMFS can be re-created on demand,
cache evictions might result in computation overhead but would not otherwise affect the functionality or correctness of the system.
In the absence of a compelling and site-specific reason,
a simple caching policy such as LRU would be sufficient.
As we will see in section ??,
LRU will also be an important piece of our final proposed approach.

@@

%Node-local scratch storage is sufficient for temporarily caching these images.
%We also assumed that durable storage is available on the bridge machine for storing the data cache and projections associated with Shrinkwrap,
%as well as a cache of container images derived from these projections.
%Again, we chose LRU as a caching policy for our evaluation,
%though administrators would be free to experiment with different policies.

%Technologies like Docker currently operate this way (often to the chagrin of users and administrators doing scientific computing).

\subsection{Problems with the initial approach}

In developing our initial approach,
we found that a number of real-world usage patterns are poorly handled.
A small change to a projection,
such as adding a library or updating dependencies,
results in the creation of a second, unrelated image.
Even a small change such as removing an unnecessary documentation file results in a completely new container image.
This is a manageable nuisance when manually testing or developing a system,
but potentially results in problematic proliferation of images in a system managing images automatically and receiving updated specifications over time from tracing jobs.

In addition to unnecessary duplication for a single user/application,
there is also significant duplication across multiple users.
In the simulation and analysis pipelines at CERN,
a number of large components such as ROOT are used nearly universally.
These components provide basic functionality and infrastructure that each experiment's software stack builds upon.
Moreover, these same common infrastructure components must be included regardless of user or application.
Thus even if every user of the system only creates a single container image,
there is a duplicate copy of all such common infrastructure components for each user.
This is problematic if we would like to design a system that can support many users at a site.

These two sources of duplication tend to compound each other,
leading us to conclude that a more sophisticated approach is necessary to work with realistically-provisioned workers and system infrastructure.
These issues are fundamentally related to unnecessary duplication of data among container images.
The same types of issues motivate the use of dynamic libraries for everyday applications.
With statically linked executables,
each application includes a complete set of required libraries.
While very stable and reliable,
OS distributions prefer to use dynamic linking.
This allows applications to share a single copy of each library,
so that foundational components such as libc are not copied into every executable,
and so that updates and changes to the set of installed packages do not require rebuilding the entire system.
For this approach to work,
package maintainers must arrange all software packages in the system to ensure compatibility,
and the OS distribution must set policies for system organization, update procedures, etc.
Container images here function similarly to statically linked executables,
but since multiple complete OS distributions are stored within CVMFS,
and since we are dealing with multi-terabyte repos and entire frameworks and software stacks rather than individual OS libraries,
the scale involved is of course larger.
In the case of CVMFS, there is no notion of a ``package mainainer'' or central authority over all software.
Each experiment manages its own software repo as it chooses,
with some components and repos shared between multiple applications.
Since we would like to be able to automatically handle existing applications that are not designed with central management in mind,
we must instead look at other approaches for dealing with data duplication.

\subsection{Deduplicating Container Images}

The representation of projections as disk images is problematic for deduplicating data.
Internally, Shrinkwrap uses hard links to a common data store so that a single file is only stored once across all projections in the system.
On creating disk images, however,
these links are broken.
Each disk image must contain a complete copy of all data,
so that a file must be included (and take up space) in every image that includes it.
Docker's solution is to build container from layers of archives that can be reused.
Docker layers are explicitly defined when writing a Dockerfile.
With CVMFS, however, there is no existing layer-like organization or universal rule that could derive layers for existing applicaions.
Working with layers defined from the outset is also a poor fit if we would like to support partial, gradually improved information from traces.
We therefore desire some sort of deduplication that can be applied for existing applications to an assortment of container images after creation.

To develop a deduplication strategy,
we started with some observations about the projection and image properties.
For ``large'' images,
there is limited opportunity for deduplication.
Each large image is a fairly complete copy of a particular repo.
While there is bound to be some overlap between versions of the same repo,
an updated image will contain a large number of updated software components,
making a ``deduplicated'' between two versions significantly larger than either individually.
There are also likely to be some conflicts in the filesystem if software components re-use the same paths for different files between versions.
This case is akin to combining RHEL~6 and RHEL~7 in an attempt to save space.
While there are similarities between the two,
the significant changes and conflicts make this a poor strategy.
Thus for our strategy we would like to avoid ``deduplicating'' images that have large non-overlapping portions.

Likewise, there is little reason to attempt to deduplicate large images built from different repos.
Such images would have little to no overlap,
so the only benefit to combining the two would be reducing the number of images by one.
This comes at the cost of needlessly storing and transferring the unused half.
This case is akin to bundling a complete Debian installation alongside RHEL.
Our deduplication strategy should only take effect for images with significant overlap.

It is also undesirable to combine images of drastically different size.
As an example, suppose a user creates a small projection as a subset of a repo for running a specific application.
This small image overlaps completely with a large image of the whole repo,
and there is likely no conflict.
Automatically combining the large and small image, however,
can have unintended consequences.
Small, tailored images minimize transfer and storage overhead,
and can be important for operating at scale.
It is therefore unwise to surprise users by suddenly making their container images several orders of magnitude larger.
This is akin to distributing a library exclusively by placing it within a disk image of a Debian installation.
We would therefore like to attempt deduplication only for
\begin{itemize}
\item similarly-sized images.
\item with large overlap.
\item and with no conflicting paths.
\end{itemize}

@@

To decide if two projections are ``close'',
we define the parameter $\alpha$ as the maximal Jaccard distance between closely related projections.
Since Jaccard distance is by definition between zero and one,
$\alpha$ must be in the same range.
This $\alpha$ parameter is something like the ``globbiness'' of the system.
Very small choices of $\alpha$ requires that projections be extremely close before considering them for merging.
In the extreme case with $\alpha = 0$,
only \emph{identical} images will be considered close so no images will be merged.
This results in a larger number of independent images.
Choosing $\alpha$ to be larger makes it more likely for images to be considered similar and merged.
This results in more augmented images that serve multiple tasks.
In the extreme case of $\alpha = 1$,
\emph{every} pair of images is considered close and merged if possible.
This results in large container images that have accumulated many projections.
We explore tradeoffs when choosing $\alpha$ and give recommendations based on observed patterns of activity in section ??.
